E1 Discrimination Analysis

### 2023-08-05 Analysis/clust_gmm.qmd
extremely minor organization

@@

lm1 <- lmer(vx~)

@@

>>>>

```{r}
#| fig.width=12
#| 
ta <- test %>% group_by(id,condit, vb) %>% summarise(mean(dist),mean(vx),mad(vx),sd(vx),IQR(vx),moments::skewness(vx))

plots <- map(names(ta)[-(1:3)], ~ {
  var_name <- .x
  ggplot(ta, aes(x = vb, y = !!sym(var_name), color = condit)) +
    gghalves::geom_half_violin()+
    gghalves::geom_half_boxplot(alpha=.5)+
    labs(x = "Velocity Band", y = var_name,
         title = paste("Plot of", var_name, "by Velocity Band and Condition"),
         color = "Condition") +
    theme_minimal()
})

# Combine the plots using patchwork
combined_plot <- wrap_plots(plots, ncol = 2)

# Print the combined plot
print(combined_plot)

dispersion_vars <- names(ta)[-(1:5)]
correlation_results <- map(dispersion_vars, ~ {
  var_name <- .x
  cor_result <- cor.test(ta[['mean(dist)']], ta[[var_name]], method = "pearson")
  list(var_name = var_name, cor_result = cor_result)
})

# Print the correlation results
correlation_results

dispersion_vars <- names(ta)[-(1:5)]
dispersion_vars <- c('mean(dist)', dispersion_vars)
ta_subset <- ta[dispersion_vars]
correlation_matrix <- cor(ta_subset, method = "pearson")
correlation_matrix

library(corrplot)

correlation_matrices <- ta %>% 
  group_by(condit) %>% 
  nest() %>% 
  mutate(correlation_matrix = map(data, ~ cor(.x[dispersion_vars], method = "pearson")))

# Create a correlation plot for each condition
correlation_plots <- correlation_matrices %>% 
  mutate(plot = map(correlation_matrix, ~ corrplot(.x, method = "color")))

# Print the correlation plots
correlation_plots$plot

GGally::ggcorr(ta[-2], aes(colour = condit), method = c("everything", "pearson"))

@@

@@

```{r}
library(EMCluster)
library(patchwork)

set.seed(1234)

x <- list(test$vx,test$vy)
out <- init.EM(x, nclass = 1)

fitK <- kmeans(test$vx,6)

ggplot(test,aes(x=vx,group=as.factor(fitK$cluster),fill=as.factor(fitK$cluster)))+geom_histogram()
ggplot(test,aes(x=vx,group=as.factor(fitK$cluster),fill=as.factor(fitK$cluster)))+geom_density()

# Decide how many clusters to look at
n_clusters <- 7

# Initialize total within sum of squares error: wss
wss <- numeric(n_clusters)

set.seed(123)

# Look over 1 to n possible clusters
for (i in 1:n_clusters) {
  # Fit the model: km.out
  km.out <- kmeans(test$vx, centers = i, nstart = 20)
  # Save the within cluster sum of squares
  wss[i] <- km.out$tot.withinss
}

# Produce a scree plot
wss_df <- tibble(clusters = 1:n_clusters, wss = wss)

scree_plot <- ggplot(wss_df, aes(x = clusters, y = wss, group = 1)) +
    geom_point(size = 4)+
    geom_line() +
    scale_x_continuous(breaks = c(2, 4, 6, 8, 10)) +
    xlab('Number of clusters')
scree_plot

```{r}
library(mixR)

mr1 <- mixfit(test$vx,ncomp=3)
plot(mr1)
mr1

mr2 <- mixfit(test$vx,ncomp=6,family="gamma")
plot(mr2)
mr2
```

library(flexmix)
#https://rdrr.io/cran/flexmix/man/flexmix.html

fm1=flexmix(vx~1|id,data=test,k=6)
summary(fm1)
plot(fm1)
parameters(fm1)

test %>% group_by(vb) %>% summarise(mean(vx),mad(vx),sd(vx),IQR(vx))

```{r}
#http://www.sthda.com/english/articles/29-cluster-validation-essentials/96-determining-the-optimal-number-of-clusters-3-must-known-methods/

e1 <- readRDS(here("data/e1_08-04-23.rds"))
test <- e1 |> filter(expMode2 == "Test") |> ungroup()
library(factoextra)
library(NbClust)

fviz_nbclust(test |> select(vx), kmeans,method = "gap_stat")

NbClust(testS |> select(vx),distance="euclidean",min.nc=2,max.nc=6,method="average")

```{r}
#| eval: false

library(brms)

testS <- test |> filter(id %in% 1:20)

options(mc.cores = 2, brms.backend = "cmdstanr")

mix <- mixture(gaussian, gaussian, gaussian,gaussian,gaussian,gaussian)
prior <- c(
  prior(normal(200, 100), Intercept, dpar = mu1),
  prior(normal(450, 100), Intercept, dpar = mu2),
  prior(normal(700, 100), Intercept, dpar = mu3),
  prior(normal(900, 100), Intercept, dpar = mu4),
  prior(normal(1100, 100), Intercept, dpar = mu5),
  prior(normal(1300, 100), Intercept, dpar = mu6)
)
fit1 <- brm(bf(vxC ~ 1), testS, family = mix,
            prior = prior, chains = 2)

summary(fit1)
pp_check(fit1,type="stat_grouped",ndraws=500,group="bandInt",stat="mean")

# saveRDS(object = fit1,
#         file = here("data/model_cache/mix6_vxC_test6.rda"), 
#         compress = "xz")

7x speedup for scaled vx
```{r}

library(brms)

test$scaleVx <- scale(test$vx)
test %>% group_by(vb) %>% summarise(mean(scaleVx))
# vb        `mean(scaleVx)`
#   <fct>               <dbl>
# 1 100-300           -0.544 
# 2 350-550           -0.273 
# 3 600-800           -0.0194
# 4 800-1000           0.449 
# 5 1000-1200          0.774 
# 6 1200-1400          1.00

testS <- test |> filter(id %in% 1:20)

options(mc.cores = 2, brms.backend = "cmdstanr")

mix <- mixture(gaussian, gaussian, gaussian,gaussian,gaussian,gaussian)
prior <- c(
  prior(normal(-.5, 10), Intercept, dpar = mu1),
  prior(normal(-.25, 10), Intercept, dpar = mu2),
  prior(normal(-.02, 10), Intercept, dpar = mu3),
  prior(normal(.4, 10), Intercept, dpar = mu4),
  prior(normal(.7, 10), Intercept, dpar = mu5),
  prior(normal(1, 10), Intercept, dpar = mu6)
)
fitScale <- brm(bf(scaleVx ~ 1), testS, family = mix,
            prior = prior, chains = 2)

summary(fitScale)
pp_check(fit1,type="stat_grouped",ndraws=500,group="bandInt",stat="mean")

# saveRDS(object = fit1,
#         file = here("data/model_cache/mix6_vxC_test6.rda"), 
#         compress = "xz")

options(mc.cores = 4, brms.backend = "cmdstanr")

mix <- mixture(gaussian, gaussian, gaussian,gaussian,gaussian,gaussian)
prior <- c(
  prior(normal(-.5, 10), Intercept, dpar = mu1),
  prior(normal(-.25, 10), Intercept, dpar = mu2),
  prior(normal(-.02, 10), Intercept, dpar = mu3),
  prior(normal(.4, 10), Intercept, dpar = mu4),
  prior(normal(.7, 10), Intercept, dpar = mu5),
  prior(normal(1, 10), Intercept, dpar = mu6)
)

fitScaleInt <- brm(bf(scaleVx ~ 1 + (1|id)), test, family = mix,
            prior = prior, chains = 4)

summary(fitScaleInt)
pp_check(fitScaleInt,type="stat_grouped",ndraws=500,stat="mean")

# saveRDS(object = fit1,
#         file = here("data/model_cache/mix6_vxC_test6.rda"), 
#         compress = "xz")

### 2023-08-05 Analysis/e1_discrim.qmd
extremely minor organization

test <- e1 |> filter(expMode2 == "Test") |> mutate(id=factor(id,levels=unique(id)))
testAvg <- test %>% group_by(id, condit, vb, bandInt,bandType,tOrder) %>%
  summarise(nHits=sum(dist==0),vxMean=mean(vx),vxMed=median(vx),dist=mean(dist),sdist=mean(sdist),n=n(),Percent_Hit=nHits/n)

@@

# fit model
m1_bayes <- brm(censored_vx | cens(censored) ~ 1 + bandInt + condit + (1 + bandInt | id), 
                data = test, family = gaussian())

m1_bayesT <- brm(censored_vx | trunc(lb = 0) ~ 1 + bandInt + condit + (1 + bandInt | id), 
                data = test, family = lognormal(),chains=1)

@@

>>>>

test <- e1 |> filter(expMode2 == "Test") 
testAvg <- test %>% group_by(id, condit, vb, bandInt,bandType,tOrder) %>%
  summarise(nHits=sum(dist==0),vxMean=mean(vx),vxMed=median(vx),dist=mean(dist),sdist=mean(sdist),n=n(),Percent_Hit=nHits/n)

@@

# fit model
m1_bayes <- brm(vx~ 1 + bandInt + condit + (1 + bandInt | id), 
                data = test, family = gaussian())

@@

$$
\begin{aligned}
  \operatorname{vx}_{i}  &\sim N \left(\alpha_{j[i]} + \beta_{1j[i]}(\operatorname{bandInt}), \sigma^2 \right) \\    
\left(
  \begin{array}{c} 
    \begin{aligned}
      &\alpha_{j} \\
      &\beta_{1j}
    \end{aligned}
  \end{array}
\right)
  &\sim N \left(
\left(
  \begin{array}{c} 
    \begin{aligned}
      &\gamma_{0}^{\alpha} + \gamma_{1}^{\alpha}(\operatorname{condit}_{\operatorname{Varied}}) \\
      &\mu_{\beta_{1j}}
    \end{aligned}
  \end{array}
\right)
, 
\left(
  \begin{array}{cc}
     \sigma^2_{\alpha_{j}} & \rho_{\alpha_{j}\beta_{1j}} \\ 
     \rho_{\beta_{1j}\alpha_{j}} & \sigma^2_{\beta_{1j}}
  \end{array}
\right)
 \right)
    \text{, for id j = 1,} \dots \text{,J}
\end{aligned}
$$

`r cat(extract_eq(m1, use_coefs = TRUE))`

$$
\begin{aligned}
  \operatorname{\widehat{vx}}_{i}  &\sim N \left(480.37_{\alpha_{j[i]}} + 0.63_{\beta_{1j[i]}}(\operatorname{bandInt}), \sigma^2 \right) \\    
\left(
  \begin{array}{c} 
    \begin{aligned}
      &\alpha_{j} \\
      &\beta_{1j}
    \end{aligned}
  \end{array}
\right)
  &\sim N \left(
\left(
  \begin{array}{c} 
    \begin{aligned}
      &56.83_{\gamma_{1}^{\alpha}}(\operatorname{condit}_{\operatorname{Varied}}) \\
      &0
    \end{aligned}
  \end{array}
\right)
, 
\left(
  \begin{array}{cc}
     351.8 & -0.8 \\ 
     -0.8 & 0.38
  \end{array}
\right)
 \right)
    \text{, for id j = 1,} \dots \text{,J}
\end{aligned}
$$
